### 2023-05-20 Documentation/UFMG/PFCII/LiteratureReview/LiteratureReview.tex
documentation

Through the various classifications and subclassifications of spraying defined in literature it will be aggregated some of them and separated between 5 modes as explained below in order of growing electric potential.

@@

Dripping mode happens when the electric field applied is not enough to change the meniscus shape, phenomenon called field enhanced dripping.
In that situation the liquid droplet has, in general, size bigger than the capillary and low frequency intervals between each drop.

@@

The Cone jet mode is a particular interest in applications, where the meniscus emits a steady microscopic jet, resulting in uniformly sized droplets. 
Taylor (1964)\cite{taylor} was the first to demonstrate that electrostatic pressure and capillary pressure can be balanced at any point on the surface of a liquid cone.
The electrified meniscus can take on a conical shape within a certain range of applied voltage and injected liquid flow rate.

In Taylor cone, here referred as Cone Jet, the direction of the cone apex also varies depending on the balance of forces. 
With a fixed flow of liquid it is noticeable that the pointing direction curves and the meniscus reduce volume when increase the electric potential until the moment the jet breaks into two or more jets going to the next classification.

@@

Zeleny's (1915)\cite{zeleny} paper was the first reference to multi-jet mode. In his experiments was observed:
\emph{"for voltages well above that for which the steady conical surface changed into a blurred agitated one, the appearance became a number of fine points with their attendant dark streaks arranged along the circumference of the raised edge".}

@@

C.N. Ryan (2012)\cite{Ryan} focused his work in multi jet spraying mode, and its current properties. 
He detected that for a low number of jets emitted the meniscus is not bound to the rim of the capillary. 
This mode is seen to be less stable, in that each jet in multi-jet mode is unable to support the same volumetric flow as in the single jet mode.
As the number of jets increases the mode becomes what Ryan refers at rim-mode electrospray. 
In this work, all modes with more than one stable jet will be classified as multi jet. Figure \ref{fig:multijt} shows three different variations of multi jet spraying mode.

@@

In this chapter it was exhibited some electrospray theory that was important for the development of this project. 
Firstly, with a brief review about the physical concept behind the experiment. 
Then, introduced how current analysis can be useful for applications citing works already done with this approach.
Finally, ending up the chapter defining what spraying modes classification will be considered and its properties.

>>>>

About the classification, there is not a convention in literature. Some authors can describe more classifications than others, hence this work will focus on 5 electrospraying modes as will be explained below.

@@

When the applied electric field is insufficient to alter the shape of the meniscus, the liquid droplet typically has a size larger than that of the capillary. Additionally, there are low-frequency intervals between each drop, indicating that the droplet is in a dripping mode.
With the increase of electric field, starts a phenomenon called field enhanced dripping, where droplet size decrease and frequency starts to increase in order to keep the same volume of liquid flowing.

@@

The Cone jet mode is a particular interest in applications, where the meniscus emits a steady microscopic jet, resulting in uniformly sized droplets. 
Taylor (1964)\cite{taylor} was the first to demonstrate that electrostatic pressure and capillary pressure can be balanced at any point on the surface of a liquid cone.
The electrified meniscus can assume a conical shape within a certain range of applied voltage and injected liquid flow rate.

In Taylor cone, here referred as Cone Jet, the direction of the cone apex also varies depending on the balance of forces. 
With a fixed flow of liquid, it is noticeable that the pointing direction curves and the meniscus reduce volume when the electric potential increase, until the moment the jet breaks into two or more jets going to the next classification.

@@

Zeleny's (1915)\cite{zeleny} paper was the first reference to multi-jet mode. In his experiments it was observed:
\emph{"for voltages well above that for which the steady conical surface changed into a blurred agitated one, the appearance became a number of fine points with their attendant dark streaks arranged along the circumference of the raised edge".}

@@

C.N. Ryan (2012)\cite{Ryan} focused his work in multi jet spraying mode, and its current properties. 
It was detected that for a low number of jets emitted the meniscus is not bound to the rim of the capillary. 
This mode is seen to be less stable, in that each jet in multi-jet mode is unable to support the same volumetric flow as in the single jet mode.
As the number of jets increases the mode becomes what Ryan refers to as rim-mode electrospray. 
In this work, all modes with more than one stable jet will be classified as multi jet. Figure \ref{fig:multijt} shows three different variations of multi jet spraying mode.

@@

This chapter presents some electrospray theory that was important for the development of this project:
Firstly, a brief review about the physical concept behind the experiment. 
Then, it is introduced how current analysis can be useful for applications, citing works already done with this approach.
Finally, ending up the chapter defining what spraying modes classification will be considered and its properties.

### 2023-05-20 Documentation/UFMG/PFCII/Methodology/Methodology.tex
documentation

\subsection{Statistical Analysis}
In Sjaaks\cite{Sjaaks} work, the author exposed signal characteristics that can be used to classify the actual spraying mode with a sample of measured current using both time and frequency domain analysis.
It is acquired the current data frame of 0.5s with 100 kHz as sampling frequency, with each current sample as 50 thousand current values. Through statistical analysis in these values such as mean, median and standard deviation the classification is done. 
Figure \ref{fig:sjaaks_statistical_class} illustrate data samples being classified using Sjaaks\cite{Sjaaks} method.

@@

    The program developed is capable of running different types of routines and with an easy way to implement new strategies.
    Continuing the methodology, in the setup json file there is a "sequence" attribute which can be chosen between "ramp", "step", "map" or "control".
    The controller thread will manage what the algorithm must do for each sequence.
    Following the control model in Figure\ref{fig:control_model_fig}, the controller outputs (voltage and flow rate) are the actuators signal.
    The sampling rate is fixed to 0.5 seconds for all the experiments, and it's managed by the data\_acquisiton\_thread\ref{subsec:data_aquisition}.

>>>>

\subsection{Statistical Analysis}
In Sjaaks\cite{Sjaaks} work, the author exposed signal characteristics that can be used to classify the actual spraying mode with a sample of measured current using both time and frequency domain analysis.
A current data frame of 0.5s is acquired with 100 kHz as sampling frequency. Therefore, each current sample holds 50 thousand current values.

Through statistical analysis in these signal such as mean, median and standard deviation the classification is done. 
Figure \ref{fig:sjaaks_statistical_class} illustrates data samples being classified using Sjaaks\cite{Sjaaks} method.

@@

    The program developed is capable of running different types of routines and with an easy way to implement new strategies.
    Continuing the methodology, in the setup json file there is a "sequence" attribute which can be chosen between "ramp", "step", "map" or "control".
    The controller thread will manage what the algorithm must do for each sequence.
    Following the control model in Figure\ref{fig:control_model_fig}, the controller outputs (voltage and flow rate) are the actuators signal.
    The sampling rate is fixed to 0.5 seconds for all the experiments, and it's managed by the data\_acquisiton\_thread.

### 2023-05-20 Documentation/UFMG/PFCII/Results/Results.tex
documentation

The automated experiment routine is capable of acquiring a significantly more precise and extensive amount of data than what can be achieved by a human. 
With the save data by streaming in saving thread \ref{subsec:save_thread}, it was achieved less program memory during the experiment and, specially, safety for not losing the data in case the program fail during an experiment.

>>>>

The automated experiment routine is capable of acquiring a significantly more precise and extensive amount of data than what can be achieved by a human. 
With the save data by streaming in saving thread, it was achieved less program memory during the experiment and, specially, safety for not losing the data in case the program fail during an experiment.

### 2023-05-20 Documentation/UFMG/PFCII/systemDescription/systemDescription.tex
documentation

In this chapter, will be discussed the equipment used in detail, including its integration and how it was modeled and implemented in both hardware and software. 
The python automation routine that is the focus of this project was started by a previous student\cite{Monica}, and the software itself was originally created as an electrospray multipurpose library.

@@

    Wireless communication allows measurements disconnected to an external power supply, which increase safety when using high voltage potential references and also reduce the signal noise collected from external power lines.
    The current is routed directly via the input, hence the oscilloscope measures the voltage dropped via its input resistance (which can be switched between 1 or 2 Mohms).
    TiePie WifiScope WS6 has a resolution of up to 16 bit at a minimal input range of 200 mV, sufficient to measure currents down to 1 nA.

    The interface with the software was made using the TiePie Library\cite{TiePieLib} and can be found in \emph{configuration\_tiepie.py}. Note that is also important to have the \emph{printinfo.py} file in the project folder in order to work.

@@

  The stability of the system is affected by many physical effects. Evidently having the more parameters analyzed favors the system control.
  The surface tension force is dependent of the liquid-gas interface on the meniscus. Hence, the surrounding gas must be constantly the same and so its humidity.
  Also, temperature is a variable that interfere in many phenomena in the system. Specially the liquid properties such as viscosity.

  For that, a standard microcontroller development board (\emph{Arduino Uno}) with a temperature and humidity sensor (DHT11) was configured to add that data in real time in the routine.
  The Arduino code can be found in the \emph{/peripherals} folder.

@@

  The pump integration in the automation algorithm was done in this work, bringing a new controllable variable, the flow rate. The spraying mode can be controlled using the two main variables that affect the system. 
  It brings more complexity for the system since now it is a multivariable control.
  Controlling also the flow rate provide this project a new dimension in the system giving freedom to explore the flow rate properties.
  With this new input variable the control system is MISO (Multiple Inputs Single Output).

  About the pump interface, as there was no good ready-to-use library, it was developed a simple and intuitive interface to be used in the software routine.
  The communication protocol used is RS-232 and the pump commands list were found in the user manual.

@@

Figure \ref{fig:setup} illustrate a diagram with all the key components in the system. The peripherals are connected to a computer running the software routine via serial communication. 
This diagram encapsulates the process system and will be used as a sub-system of the control model. The input in the process are power supply voltage and pump machine flow rate, referred here as the controller values. The output is the oscilloscope current sample.

@@

There are also many minor variables that affects the experiment stability. More details are explained in appendix \ref{sec:setup_validation}.

@@

The software was reformulated on top of the control model represented in Figure \ref{fig:control_model_fig}. The process of the control loop is the same as represented in Figure \ref{fig:setup}. 
Each other subsystem in the control loop is a separate thread that will be explained bellow.

@@

    In order to implement this closed loop control model to the software and explore parallel processing, each sub-system was implemented as a separate thread.
    For concurrency on flux of data between threads was used queues structures.
    A queue is an abstract data type that holds an ordered, linear sequence of items. You can describe it as a first in, first out (FIFO) structure.

    \subsection{Controller Thread}

        The controller thread is responsible for sending the power supply set voltage values and the syringe pump the flow rate set values according to the sequence selected.
        Also, responsible for sending the finish event command that end the routine and trigger the threads to close their routines.
        As input, there is the setup config file and the \emph{feedback\_queue}. As output, the values in the \emph{controller\_output\_queue()}.

    \subsection{Data Acquisition Thread}
    \label{subsec:data_aquisition}

        The data acquisition thread is responsible for reading the current data from the oscilloscope, humidity and temperature data from the DHT11 sensor, voltage from the power supply, flow rate from the pump and concatenate into one sample data.
        As output, the values in the \emph{data\_queue()}.

    \subsection{Data Processing Thread}

        The data processing is responsible for calculating the statistical values from the raw data and classify it in the respective spray mode for that sample.
        As output, the values in the \emph{save\_data\_queue()}, \emph{plotting\_queue()} and \emph{feedback\_queue()}.

    \subsection{Save Data Thread}
    \label{subsec:save_thread}

        After processing, the data is saved in real time in a json file using \emph{jsonstreams} library to save one sample structure at a time.
        With the new streaming model of saving, instead of having all data together in program memory to be saved in the end, it is saving each sample in real time during the experiment.
        The data acquired in each sample of 0.5s is shown in Figure \ref{fig:data_sample}.

@@

        This new way of saving the measurement and processing data altogether separate by enumerated samples prevents loss of data in case of error during experiments and, specially improve in the usability of the data for further deep analysis. One experiment of around 1 hour collects around 6 GB of data. Organizing that amount of data to make it fast and easy to use was an important improvement in the software.
        To work and analyze the data it can be used pandas Data frame. A python widely used framework for data analysis.
        With the command:

@@

        To conclude, json files are good as a system output because of its human readability. But as the database gets bigger json becomes to be a slow to read and heavy to store data. For that, saving the data frame in a compressed type of file called feather is much faster to work with it.

    \subsection{Plotting}

        The only running function that should run in the main thread because the plotting library \emph{matplotlib} incompatibilities of running outside the main function. 
        The plotting function is responsible for plotting in real time the current sample acquired, and its respective fast Fourier transform to evaluate the sample frequency spectrum.
        It takes as input the \emph{plot\_data\_queue} from the \emph{processing\_thread()} and displays three graphs updated in real time on the screen during the experiment.

@@

    In this chapter it is described the system description detailing the hardware instruments, how it was modelled and its implementation in the software.

>>>>

In this chapter, the equipment used will be discussed in detail, including its integration and how it was modeled and implemented in both hardware and software. 
The python automation routine that is the focus of this project was started by a previous student\cite{Monica}, and the software itself was originally created as an electrospray multipurpose library.

@@

    Wireless communication allows measurements disconnected from an external power supply, which increases safety when using high voltage potential references and also reduce the signal noise collected from external power lines.
    The current is routed directly via the input, hence the oscilloscope measures the voltage dropped via its input resistance (which can be switched between 1 or 2 Mohms).
    TiePie WifiScope WS6 has a resolution of up to 16 bit at a minimal input range of 200 mV, sufficient to measure currents down to 1 nA. In addition the WifiScope has an open source interface library for python, \cite{TiePieLib}, making it easier to integrate with the automation routine.

@@

  The stability of the system is affected by many physical effects. Evidently, it favors the system control having more parameters analyzed.
  The surface tension force, for example, is dependent of the liquid-gas interface on the meniscus, hence, the surrounding gas must be constant and so its humidity.
  Also, temperature is a variable that interferes in many phenomena in the system, specially the liquid properties such as viscosity.

  For that, a standard microcontroller development board (\emph{Arduino Uno}) with a temperature and humidity sensor (DHT11) was configured to add that data in real time to the routine.

@@

  The pump integration in the automation algorithm was done in this work, bringing a new controllable variable, the flow rate. The spraying mode can be controlled using the two main variables that affect the system. 
  It brings more complexity for the system since now it is a multivariable control.
  Controlling also the flow rate provides this project a new dimension in the system giving freedom to explore the flow rate properties.
  With this new input variable the control system is MISO (Multiple Inputs Single Output).

  About the pump interface, as there was no good ready-to-use library, a simple and intuitive interface was developed to be used in the software routine.
  The communication protocol used is RS-232 and the pump commands list were found in the user manual.

@@

Figure \ref{fig:setup} illustrates a diagram with all the key components in the system. The peripherals are connected to a computer running the software routine via serial communication. 
This diagram encapsulates the process system and will be used as a sub-system of the control model. The inputs in the process are power supply voltage and pump machine flow rate, referred here as the controller values. The output is the oscilloscope current measurement.

@@

There are also many minor variables that affect the experiment stability. More details are explained in appendix \ref{sec:setup_validation}.

@@

The software was reformulated on top of the control model represented in Figure \ref{fig:control_model_fig}. The process of the control loop is the same as represented in Figure \ref{fig:setup}. 
Each other subsystem in the control loop is a separate thread that will be explained below.

@@

    In order to implement this closed loop control model in the software and explore parallel processing, each sub-system was implemented as a separate thread.
    For concurrency on flux of data between threads was used queues structures.
    A queue is an abstract data type that holds an ordered, linear sequence of items. You can describe it as a first in, first out (FIFO) structure.

    The controller thread is responsible for sending to the power supply the set voltage values and to syringe pump the flow rate set values according to the sequence selected.
    Also, responsible for sending the finish event command that end the routine and trigger the threads to close their routines.
    As input, the setup config file with information about the liquid and the setup used such as geometric parameters of the nozzle to plate system. As output, the commands to the actuators.

    The data acquisition thread is responsible for concatenating the current, voltage, flow rate, humidity and temperature data and concatenate it into one sample.
    The data processing is responsible for calculating the statistical values from the raw data and classify it in the respective spray mode for that sample.
    After processing, the data is saved in real time in a json file using \emph{jsonstreams} library to save one sample structure at a time.
    With the new streaming model of saving, instead of having all data together in program memory to be saved in the end, it is saving each sample in real time during the experiment.
    The data acquired in each sample of 0.5s is shown in Figure \ref{fig:data_sample}.

@@

    This new way of saving the measurement and processing ensures that the data is stored separately in enumerated samples.
    This prevents any loss of data in the event of errors during experiments and significantly enhances the usability of the data for in-depth analysis. Each hour-long experiment generates approximately 6 GB of data, and effectively organizing such a substantial amount of data was a crucial enhancement in the software. To efficiently work with and analyze the data, the widely-used Python framework, pandas Dataframe, can be utilized with the following command:

@@

    To conclude, json files are good as a system output because of its human readability. But as the database gets bigger json becomes slow to read and large to store. For that, saving the data frame in a compressed type of file called feather is much faster to work with it.

    This is only running function that should run in the main thread because of the plotting library \emph{matplotlib} incompatibilities with running outside the main function. 
    The plotting function is responsible for plotting in real time the current sample acquired, and its respective fast Fourier transform to evaluate the sample frequency spectrum.
    It takes as input the \emph{plot\_data\_queue} from the \emph{processing\_thread()} and displays three graphs updated in real time on the screen during the experiment.

@@

    In this chapter it is explained the system description, detailing the hardware instruments, how it was modelled and its implementation in the software.
